---
title: "finalprojstart"
output: html_document
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  warning = FALSE,
  message = FALSE,
  echo = TRUE
)
q <- 0
```

```{r setup, include=FALSE}
library(tidyverse)
library(tidycensus)
library(patchwork)
library(ggthemes)
library(rvest)
library(janitor)
library(lubridate)
library(httr2)
library(spotifyr)
# library(devtools)
# install_github("charlie86/spotifyr")
library(purrr)
```

```{r}
client_id <- readLines("spotify_client_id.txt")
client_secret <- readLines("spotify_client_secret.txt")

get_spotify_token <- function(client_id, client_secret) {
  resp <- request("https://accounts.spotify.com/api/token") |>
    req_auth_basic(client_id, client_secret) |>
    req_body_form(grant_type = "client_credentials") |>
    req_method("POST") |>
    req_perform()
  
  resp_body_json(resp)$access_token
}

token <- get_spotify_token(client_id, client_secret)
```

# 2024 Top 10 Albums
```{r}
# Function to get album data
get_albums_data <- function(album_ids) {
  req <- request("https://api.spotify.com/v1/albums") |>
    req_url_query(ids = paste(album_ids, collapse = ",")) |>
    req_headers(Authorization = paste("Bearer", token))

  resp <- req_perform(req)

  albums <- resp |>
    resp_body_json(simplifyVector = FALSE) |>
    (\(x) x$albums)()

  map_dfr(albums, function(album) {
    tibble(
      album_id = album$id,
      album_name = album$name,
      artist_id = album$artists[[1]]$id,
      artist_name = album$artists[[1]]$name,
      release_date = ymd(album$release_date),
      total_tracks = album$total_tracks,
      genre = paste(artist$genres, collapse = ", ")
    )
  })
}

# Function to get genres for each artist
get_artist_genres <- function(artist_id) {
  req <- request(paste0("https://api.spotify.com/v1/artists/", artist_id)) |>
    req_headers(Authorization = paste("Bearer", token))

  resp <- req_perform(req)

  genres <- resp |>
    resp_body_json(simplifyVector = TRUE) |>
    (\(x) x$genres)()

  return(paste(genres, collapse = ", "))
}

# Add genres to album data
album_data2 <- album_data |>
  rowwise() |>
  mutate(genres = get_artist_genres(artist_id)) |>
  ungroup()

album_ids <- c(
  "5H7ixXZfsNMGbIE5OBSpcb",
  "7aJuG4TFXa2hmE4z1yxc3n",
  "3iPSVi54hsacKKl1xIR2eH",
  "4kS7bSuU0Jm9LYMosFU2x5",
  "5EYKrEDnKhhcNxGedaRQeK",
  "64LU4c1nfjz1t4VnGhagcg",
  "07w0rG5TETcyihsEIZR3qG",
  "1NAmidJlEaVgA3MpcPFYGq",
  "168CdR21lfn0TTyw1Pkdcm",
  "2ODvWsOgouMbaA5xf0RkJe")
album_data <- get_albums_data(album_ids)
```

# How have album release patterns changed over recent years?
```{r}
album_data %>%
  mutate(year = year(release_date), month = month(release_date, label = TRUE)) %>%
  ggplot(aes(x = factor(year))) +
  geom_bar() +
  labs(title = "Number of Album Releases by Year", x = "Year", y = "Number of Albums")
```

# What genres dominate the top albums and playlists of 2024?
```{r}
album_data2 %>%
  separate_rows(genres, sep = ", ") %>%
  count(genres, sort = TRUE) %>%
  slice_head(n = 10) %>%
  ggplot(aes(x = reorder(genres, n), y = n)) +
  geom_col(fill = "steelblue") +
  coord_flip() +
  labs(title = "Top Genres in 2024 Top Albums", x = "Genre", y = "Count")
```


# Search by Album Name (1 at a time)
```{r}
query <- "1989"

search_request <- request("https://api.spotify.com/v1/search") |>
  req_url_query(q = query, type = "album", limit = 5) |>
  req_headers(Authorization = paste("Bearer", token))

search_response <- req_perform(search_request)

results <- search_response |>
  resp_body_json(simplifyVector = TRUE)

albums <- results$albums$items

# Extract album IDs and names
album_tbl <- tibble::tibble(
  id = albums$id,
  name = albums$name,
  artist = sapply(albums$artists, \(a) paste(a$name, collapse = ", ")),
  release_date = albums$release_date
)

print(album_tbl)
```

# Search Multiple Albums
```{r}
queries <- c("1989", "Short n’ Sweet")

get_albums <- function(query, token) {
  search_request <- request("https://api.spotify.com/v1/search") |>
    req_url_query(q = query, type = "album", limit = 5) |>
    req_headers(Authorization = paste("Bearer", token))
  
  search_response <- req_perform(search_request)
  results <- search_response |>
    resp_body_json(simplifyVector = TRUE)
  
  albums <- results$albums$items
  
  tibble(
    id = albums$id,
    name = albums$name,
    artist = sapply(albums$artists, \(a) paste(a$name, collapse = ", ")),
    release_date = albums$release_date,
    query = query
  )
}

album_tbl <- map_dfr(queries, ~ get_albums(.x, token))

print(album_tbl)
```

# Plot Album Name by Release Year
```{r}
# Ensure release_date is a date object
album_tbl$release_date <- ymd(album_tbl$release_date)

# Extract year
album_tbl$release_year <- year(album_tbl$release_date)

# Plot
ggplot(album_tbl, aes(x = factor(release_year), fill = artist)) +
  geom_bar() +
  labs(
    title = "Albums Named '1989' by Release Year",
    x = "Release Year",
    y = "Number of Albums",
    fill = "Artist"
  ) +
  theme_minimal()
```


# "Which artists have the most tracks in top playlists, and how do their popularity and follower counts compare?"
```{r}
# Count tracks per artist
top_artists <- playlist_tracks %>%
  count(artist_name, artist_id, sort = TRUE)

# Plot top 10 artists by number of tracks
top_artists %>%
  slice_max(n, n = 10) %>%
  ggplot(aes(x = reorder(artist_name, n), y = n, fill = artist_name)) +
  geom_col(show.legend = FALSE) +
  coord_flip() +
  labs(
    title = "Top 10 Artists by Track Count in Playlist",
    x = "Artist",
    y = "Number of Tracks"
  ) +
  theme_minimal()
```



# Plots for the 2024 Top Songs Playlist
```{r}
# Histogram of track durations in minutes
playlist_tracks %>%
  mutate(duration_min = track_duration_ms / 60000) %>%
  ggplot(aes(x = duration_min)) +
  geom_histogram(binwidth = 0.5, fill = "steelblue", color = "black") +
  labs(
    title = "Distribution of Track Durations in Playlist",
    x = "Duration (minutes)",
    y = "Number of Tracks"
  ) +
  theme_minimal()

# Bar plot of explicit vs non-explicit tracks
playlist_tracks %>%
  mutate(explicit = ifelse(track_explicit, "Explicit", "Not Explicit")) %>%
  count(explicit) %>%
  ggplot(aes(x = explicit, y = n, fill = explicit)) +
  geom_col() +
  labs(
    title = "Explicit Content in Playlist Tracks",
    x = "Explicit",
    y = "Count of Tracks"
  ) +
  scale_fill_manual(values = c("tomato", "forestgreen")) +
  theme_minimal()

# Top 10 artists by number of tracks in playlist
playlist_tracks %>%
  count(artist_name, sort = TRUE) %>%
  slice_head(n = 10) %>%
  ggplot(aes(x = reorder(artist_name, n), y = n, fill = artist_name)) +
  geom_col(show.legend = FALSE) +
  coord_flip() +
  labs(
    title = "Top 10 Artists by Number of Tracks in Playlist",
    x = "Artist",
    y = "Number of Tracks"
  ) +
  theme_minimal()
```


```{r}
get_artist_genres <- function(artist_ids, token) {
  split_ids <- split(artist_ids, ceiling(seq_along(artist_ids) / 50))
  
  map_dfr(split_ids, function(ids_chunk) {
    req <- request("https://api.spotify.com/v1/artists") |>
      req_url_query(ids = paste(ids_chunk, collapse = ",")) |>
      req_headers(Authorization = paste("Bearer", token))
    
    resp <- req_perform(req)
    content <- resp_body_json(resp, simplifyVector = FALSE)
    
    map_dfr(content$artists, function(artist) {
      tibble(
        artist_id = artist$id,
        artist_name = artist$name,
        genres = paste(artist$genres, collapse = ", ")
      )
    })
  })
}
```

# Top 2024 Songs Playlist with Genres Column Added
```{r}
# 1. Get unique artist IDs from tracks
artist_ids <- unique(playlist_tracks$artist_id)

# 2. Get genres for those artists
artist_genres <- get_artist_genres(artist_ids, token)

# 3. Join genres back to tracks by artist_id
tracks_with_genres <- playlist_tracks |>
  left_join(artist_genres, by = "artist_id")
```

# Top Songs Of 1 Artist (with Popularity rating)
```{r}
get_artist_top_tracks <- function(artist_id, token, market = "US") {
  url <- paste0("https://api.spotify.com/v1/artists/", artist_id, "/top-tracks")
  
  req <- request(url) |>
    req_url_query(market = market) |>
    req_headers(Authorization = paste("Bearer", token))
  
  resp <- req_perform(req)
  
  content <- resp_body_json(resp, simplifyVector = FALSE)
  
  tracks <- content$tracks
  
  map_dfr(tracks, function(track) {
    tibble(
      track_id = track$id,
      track_name = track$name,
      album_id = track$album$id,
      album_name = track$album$name,
      artist_id = track$artists[[1]]$id,
      artist_name = track$artists[[1]]$name,
      track_duration_ms = track$duration_ms,
      track_explicit = track$explicit,
      popularity = track$popularity,
      preview_url = track$preview_url
    )
  })
}

# Usage example:
artist_id <- "06HL4z0CvFAxyc27GXpf02"  # Taylor Swift's artist ID
# token <- "your_access_token_here"
top_tracks <- get_artist_top_tracks(artist_id, token, market = "US")
```

# Top 10 Artists + Genre + Popularity
```{r}
get_artists_data <- function(artist_ids, token) {
  # Spotify API allows max 50 artist IDs per request, so split if needed
  artist_chunks <- split(artist_ids, ceiling(seq_along(artist_ids) / 50))
  
  map_dfr(artist_chunks, function(ids_chunk) {
    req <- request("https://api.spotify.com/v1/artists") |>
      req_url_query(ids = paste(ids_chunk, collapse = ",")) |>
      req_headers(Authorization = paste("Bearer", token))
    
    resp <- req_perform(req)
    content <- resp_body_json(resp, simplifyVector = FALSE)
    
    artists <- content$artists
    
    map_dfr(artists, function(artist) {
      tibble(
        artist_id = artist$id,
        artist_name = artist$name,
        genres = paste(artist$genres, collapse = ", "),
        popularity = artist$popularity,
        followers = artist$followers$total,
        artist_url = artist$external_urls$spotify
      )
    })
  })
}

# Usage example:
artist_ids <- c("06HL4z0CvFAxyc27GXpf02", "1Xyo4u8uXC1ZmMpatF05PJ", "4q3ewBCX7sLwd24euuV69X", "3TVXtAsR1Inumwj472S9r4", "6qqNVTkY8uBg9cP3Jd7DAH", "0Y5tJX1MQlPlqiwlOH1tJY", "12GqGscKJx3aE4t07u7eVZ", "5K4W6rqBFWDnAN6FQUkS6x", "66CXWjxzNUsdJxJ2JdwvnR", "2LRoIwlKmHjgvigdNGBHNo") 
# token <- "your_access_token_here"
artists_data <- get_artists_data(artist_ids, token)
```

```{r}
# Merge artist metadata with track count
artist_summary <- top_artists %>%
  left_join(artists_data, by = "artist_id")

# Scatter plot
ggplot(artist_summary, aes(x = followers, y = popularity, size = n, color = artist_name.x)) +
  geom_point(alpha = 0.8, show.legend = FALSE) +
  scale_x_log10(labels = scales::label_comma()) +
  labs(
    title = "Artist Popularity vs Followers",
    subtitle = "Bubble size = Number of Tracks in Playlist",
    x = "Followers (log scale)",
    y = "Popularity"
  ) +
  theme_minimal()
```




