---
title: "Spotify Top Artists/Songs of 2024 Globally"
output: html_document
---

```{r setup, include=FALSE}
library(tidyverse)
library(httr2)
library(plotly)
library(DT)
```

```{r}
client_id <- readLines("spotify_client_id.txt")
client_secret <- readLines("spotify_client_secret.txt")

get_spotify_token <- function(client_id, client_secret) {
  resp <- request("https://accounts.spotify.com/api/token") |>
    req_auth_basic(client_id, client_secret) |>
    req_body_form(grant_type = "client_credentials") |>
    req_method("POST") |>
    req_perform()
  
  resp_body_json(resp)$access_token
}

token <- get_spotify_token(client_id, client_secret)
```

```{r}
get_artists_data <- function(artist_ids, token) {
  artist_chunks <- split(artist_ids, ceiling(seq_along(artist_ids) / 50))
  
  map_dfr(artist_chunks, function(ids_chunk) {
    req <- request("https://api.spotify.com/v1/artists") |>
      req_url_query(ids = paste(ids_chunk, collapse = ",")) |>
      req_headers(Authorization = paste("Bearer", token))
    
    resp <- req_perform(req)
    content <- resp_body_json(resp, simplifyVector = FALSE)
    
    map_dfr(content$artists, function(artist) {
      tibble(
        artist_id = artist$id,
        artist_name = artist$name,
        genres = paste(artist$genres, collapse = ", "),
        popularity = artist$popularity,
        followers = artist$followers$total,
        artist_url = artist$external_urls$spotify
      )
    })
  })
}

get_artist_top_tracks <- function(artist_id, token) {
  url <- paste0("https://api.spotify.com/v1/artists/", artist_id, "/top-tracks?market=US")
  
  req <- request(url) |>
    req_headers(Authorization = paste("Bearer", token))
  
  resp <- req_perform(req)
  content <- resp_body_json(resp, simplifyVector = FALSE)
  
  map_dfr(content$tracks, function(track) {
    tibble(
      track_name = track$name,
      album_name = track$album$name,
      popularity = track$popularity,
      track_duration_ms = track$duration_ms,
      explicit = track$explicit,
      artist_id = artist_id
    )
  })
}

artist_ids_2024 <- c(
  "06HL4z0CvFAxyc27GXpf02",  # Taylor Swift
  "1Xyo4u8uXC1ZmMpatF05PJ",  # The Weeknd
  "4q3ewBCX7sLwd24euuV69X",  # Bad Bunny
  "3TVXtAsR1Inumwj472S9r4",  # Drake
  "6qqNVTkY8uBg9cP3Jd7DAH",  # Billie Eilish
  "0Y5tJX1MQlPlqiwlOH1tJY",  # Travis Scott
  "12GqGscKJx3aE4t07u7eVZ",  # Peso Pluma
  "5K4W6rqBFWDnAN6FQUkS6x",  # Kanye West
  "66CXWjxzNUsdJxJ2JdwvnR",  # Ariana Grande
  "2LRoIwlKmHjgvigdNGBHNo"   # Feid
)

artists_data_2024 <- get_artists_data(artist_ids_2024, token)
all_tracks_2024 <- map_dfr(artist_ids_2024, get_artist_top_tracks, token = token)

# Merge track data with artist names
all_tracks_2024 <- left_join(all_tracks_2024, artists_data_2024 |> select(artist_id, artist_name), by = "artist_id")
```


## Top 10 Most Popular Songs (All Artists)

```{r}
top_tracks <- all_tracks_2024 |> 
  arrange(desc(popularity))

datatable(top_tracks |> 
            select(Track = track_name,
                   Artist = artist_name,
                   Popularity = popularity,
                   Explicit = explicit))
```

## Popularity by Track Name (Color by Explicit)

```{r}
all_tracks_2024 |>
  filter(artist_name == "Taylor Swift") |>
  plot_ly(
        x = ~reorder(track_name, popularity),
        y = ~popularity,
        type = 'bar',
        color = ~explicit,
        colors = c('blue', 'red'),
        text = ~paste("Track:", track_name,
                      "\nArtist:", artist_name,
                      "\nPopularity:", popularity,
                      "\nExplicit:", explicit),
        hoverinfo = "text") |>
  layout(title = "Top Tracks by Popularity",
         xaxis = list(title = "Track Name", tickangle = -45),
         yaxis = list(title = "Popularity"))
```

## Research Questions:
1. What are the most popular songs from each top artist?
2. Do explicit tracks tend to have higher or lower popularity scores than non-explicit tracks?
3. Is there a relationship between the number of tracks in an album and its popularity score?
4. Does the month that an album is released impact their popularity?
5. What are the most popular songs from each top album?

## Resources:
1. Revealed: The Top Artists, Songs, Albums, Podcasts, and Audiobooks of 2024 (https://newsroom.spotify.com/2024-12-04/top-songs-artists-podcasts-audiobooks-albums-trends-2024/)

2. Spotify Web API
(https://developer.spotify.com/documentation/web-api)
